Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

In addition to standard quantified formulas $\forall \vec x.\; \varphi$,
we consider \emph{annotated quantified formulas} of the form
$\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
an uninterpreted function symbol, 
whose semantics is the same as standard quantified formulas.
Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
an annotated
quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
$x_1, \ldots, x_n$ of respective type
$\tau_1, \ldots, \tau_n$ and $\varphi$ is a quantifier-free formula 
$T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
%\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
%I don't think it makes much sense to have any quantifiers within function bodies, 
%since they aren't operational, that's why I put quantifier-free. --ajr
We write $\exists
\vec x.\; \varphi$ as a shorthand for $\neg \forall \vec x.\; \neg \varphi$.

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \}$ if it is of the form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
%
where $\con{f}_1, \ldots, \con{f}_n \in \sfuns{\Sigma}$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{conjecture} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

Given a signature $\Sigma$,
we distinguish the subset of (\emph{recursively}) \emph{defined} function
symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
We consider $\Sigma$-formulas that are in definitional form with respect to
$\sfundefs{\Sigma}$.

\begin{definition}\rm
Given a set of function definitions 
$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
a ground formula $\varphi$ 
%\ct{I suppose $\varphi$ is meant to be ground.} yes --ajr
is \emph{closed under function expansion with respect to $\DDD$} if 
\[
 \varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \} .
\]
The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
$\varphi$ closed under function expansion wrt $\DDD$
the formula
$\varphi \wedge \forallf{\con{f}_1} \vec x.\; \varphi_1 \wedge \cdots \wedge \forallf{\con{f}_n} \vec x.\; \varphi_n$ is also $T$-satisfiable.
\ct{I}
\end{definition}

>>>>

In addition to standard quantified formulas $\forall \vec x.\; \varphi$,
we consider \emph{annotated quantified formulas} of the form
$\forallf{\const{f}} \vec x.\; \varphi$, where $\con{f} \in \sfuns{\Sigma}$ is
an uninterpreted function symbol, 
whose semantics is the same as standard quantified formulas.
Given $\con{f} : \tau_1 \times \cdots \times \tau_n \rightarrow \tau$,
an annotated
quantified formula $\forallf{\con{f}} \vec x.\; \varphi$ is a \emph{function definition}
(\,\emph{for $\con{f}$}\vthinspace) if $\vec x$ is a tuple of variables
$x_1 : \tau_1$, $\ldots,$ $x_n : \tau_n$
and $\varphi$ is a quantifier-free formula 
$T$-equivalent to $\con{f}( \vec x ) \teq t$ for some term $t$ of type $\tau$.
%\ct{Why can't $\varphi$ be existential instead of just quantifier free?} 
%I don't think it makes much sense to have any quantifiers within function bodies, 
%since they aren't operational, that's why I put quantifier-free. --ajr
We write $\exists
\vec x.\; \varphi$ as a shorthand for $\neg \forall \vec x.\; \neg \varphi$.

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
\[(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi\]
%
where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
\jb{do we?}
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{conjecture} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

Given a signature $\Sigma$, we distinguish the subset of \emph{defined}
function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
We consider $\Sigma$-formulas that are in definitional form with respect to
$\sfundefs{\Sigma}$.

\begin{definition}\rm
Given a set of function definitions 
$\DDD = \{ \forallf{\con{f}_1} \vec x.\; \varphi_1, \ldots, \forallf{\con{f}_n} \vec x.\; \varphi_n \}$, 
a ground formula $\varphi$ 
%\ct{I suppose $\varphi$ is meant to be ground.} yes --ajr
is \emph{closed under function expansion with respect to $\DDD$} if 
\[
 \varphi \,\models_T\, \bigwedge\nolimits_{\,i=1}^{n} \{ \varphi_i[ \vec t / \vec x ] \mid \con{f}_i( \vec t ) \in \terms( \varphi ) \} .
\]
The set $\DDD$ is \emph{admissible} if for every $T$-satisfiable formula 
$\varphi$ closed under function expansion with respect to $\DDD$,
the formula
$\varphi \wedge (\forallf{\con{f}_1} \vec x.\; \varphi_1) \wedge \cdots \wedge (\forallf{\con{f}_n} \vec x.\; \varphi_n)$ is also $T$-satisfiable.
\ct{I}
\end{definition}
